---
title: 'ExercisesRmd'
output: html_document
date: '2023-06-06'
---

# Intro

## Packages

Packages to install

```{r, eval=FALSE}
devtools::install_github("MiriamLL/sula")
devtools::install_github("MiriamLL/GermanNorthSea")
devtools::install_github("MiriamLL/seamonas")
install.packages('sf')
install.packages('ggspatial')
install.packages('plotly')
```

Call packages

```{r}
library(tidyverse)
library(palmerpenguins)
library(ggplot2)
library(sula)
library(GermanNorthSea)
library(seamonas)
library(ggspatial)
library(sf)
```

# ~~~~~~~~~~~~
# First part

## Data

Load data to environment

```{r}
penguins<-penguins
```

Using ggplot

```{r}
ggplot()
```

##Add data to ggplot

```{r}
ggplot(data=penguins)
```

## Add axis

```{r}
ggplot(data=penguins,
       mapping=aes(x=bill_length_mm,
                   y=bill_depth_mm))
```

## Add geometry

```{r}
ggplot(data=penguins,
       mapping=aes(x=bill_length_mm,
                   y=bill_depth_mm))+
       geom_point()
```


## Simplify 

We can skip writing data= and mapping= by just following the order of the elements.

Simplified example.

```{r}
ggplot(penguins, 
       aes(x=bill_length_mm,
           y=flipper_length_mm))+
  geom_point()
```

# ~~~~~~~~~~~~~~~~
# Second Part

## Scatterplot

```{r}
ggplot(data=penguins,mapping=aes(x=bill_length_mm,y=bill_depth_mm))+
  geom_point()
```

## Line chart

```{r}
ggplot(data=penguins,mapping=aes(x=bill_length_mm,y=bill_depth_mm))+
  geom_line()
```

## Barplot

```{r}
ggplot(data=penguins,mapping=aes(x=bill_length_mm,y=bill_depth_mm))+
  geom_bar()
```

```{r}
ggplot(data=penguins,mapping=aes(x=bill_length_mm,y=bill_depth_mm))+
  geom_bar(stat='identity')
```

## Boxplot

```{r}
ggplot(data=penguins,
       mapping=aes(x=bill_length_mm,y=bill_depth_mm))+
  geom_boxplot() 
```

```{r}
ggplot(data=penguins,
       mapping=aes(x=bill_length_mm,y=bill_depth_mm, 
                   group=species))+
  geom_boxplot()
```

## Trajectory

```{r}
tracking_data<-sula::GPS_raw
```

```{r}
ggplot(data = tracking_data, aes(x = Latitude, y = Longitude)) +
  geom_path()
```

## Map

```{r}
library(GermanNorthSea)
```

```{r}
Europe<-German_land
```

```{r}
ggplot(data=Europe)+ 
  geom_sf()
```

# ~~~~~~~~~~~~~~~~
# Third Part

# Maps

Identify directory

```{r, message=TRUE, message=FALSE}
My_directory<-here('Downloads')
list.files(My_directory)
```

```{r, message=TRUE, message=FALSE}
My_shapefile<-paste0(My_directory,"/Europe.shp")
```

Load shapefile into your environment

```{r}
library(sf)
```

```{r}
Europe_shp<-st_read(My_shapefile)
```

```{r}
ggplot(data = Europe_shp)+  
  geom_sf()
```

## Coords

```{r}
ggplot(data = Europe_shp)+  
  geom_sf()+
  coord_sf(xlim = c(6, 10),ylim = c(53, 56))
```

## Legend

```{r}
ggplot(data = Europe_shp)+  
  geom_sf()+
  coord_sf(xlim = c(6, 10),ylim = c(53, 56))+
    scale_x_continuous(labels = function(x) paste0(x, '\u00B0', "W")) +
    scale_y_continuous(labels = function(x) paste0(x, '\u00B0', "N"))
```

## Scale

```{r}
library(ggspatial)
```

```{r}
ggplot(data = Europe_shp)+  
  geom_sf()+
  coord_sf(xlim = c(6, 10),ylim = c(53, 56))+
    scale_x_continuous(labels = function(x) paste0(x, '\u00B0', "W"))+
    scale_y_continuous(labels = function(x) paste0(x, '\u00B0', "N"))+
    annotation_scale(location = "br")
```

## Arrow

```{r}
ggplot(data = Europe_shp)+  
  geom_sf()+
  coord_sf(xlim = c(6, 10),ylim = c(53, 56))+
    scale_x_continuous(labels = function(x) paste0(x, '\u00B0', "W"))+
    scale_y_continuous(labels = function(x) paste0(x, '\u00B0', "N"))+
    annotation_scale(location = "br")+
    annotation_north_arrow(location = "tl")
```

## Points

Load library

```{r}
library(seamonas)
```

Add data to environment

```{r}
survey_4326<-survey_4326
```

Plot data points

```{r}
ggplot(data = Europe_shp)+  
  geom_sf()+
  coord_sf(xlim = c(6, 10),ylim = c(53, 56))+
    scale_x_continuous(labels = function(x) paste0(x, '\u00B0', "W"))+
    scale_y_continuous(labels = function(x) paste0(x, '\u00B0', "N"))+
    annotation_scale(location = "br")+
    annotation_north_arrow(location = "tl") 
```

## Change CRS

```{r}
Europe_3035<-sf::st_transform(Europe_shp, 3035)
```

```{r}
ggplot(data=Europe_3035)+ 
  geom_sf()+
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000))
```

## Simplify

```{r}
My_plot<-ggplot(data = Europe_shp)+  
  geom_sf()+
  coord_sf(xlim = c(4, 12),ylim = c(53, 56))+
    scale_x_continuous(labels = function(x) paste0(x, '\u00B0', "W"))+
    scale_y_continuous(labels = function(x) paste0(x, '\u00B0', "N"))
```

```{r}
My_plot
```

# ~~~~~~~~~~~
# Fourth part

# Aesthetics


# ~~~~~~~~~~~~
# Fifth part

# Add geometries

# ~~~~~~~~~~~~
# Sixth part

# Export

## ggsave

# End of document

