---
title: 'ExercisesRmd'
output: html_document
date: '2023-06-06'
---

# Intro

## Packages

Packages to install

```{r, eval=FALSE}
devtools::install_github("MiriamLL/sula")
devtools::install_github("MiriamLL/GermanNorthSea")
devtools::install_github("MiriamLL/seamonas")
install.packages('sf')
install.packages('ggspatial')
install.packages('plotly')
```

Check that you have the packages

```{r}
library(tidyverse)
library(palmerpenguins)
library(ggplot2)
library(sula)
library(GermanNorthSea)
library(seamonas)
library(ggspatial)
library(sf)
```

# ~~~~~~~~~~~~
# First part
# Theory

## Data

Load data to environment

```{r}
penguins<-penguins
```

Using ggplot

##Add data to ggplot

```{r}
ggplot(data=penguins)
```

## Add axis

```{r}
ggplot(data=penguins,
       mapping=aes(x=bill_length_mm,
                   y=bill_depth_mm))
```

## Add geometry

```{r}
ggplot(data=penguins,
       mapping=aes(x=bill_length_mm,
                   y=bill_depth_mm))+
       geom_point()
```


## Simplify 

We can skip writing data= and mapping= by just following the order of the elements.

Simplified example.

```{r}
ggplot(penguins, 
       aes(x=bill_length_mm,
           y=flipper_length_mm))+
  geom_point()
```

# ~~~~~~~~~~~~~~~~
# Second Part
# Geometries

## Scatterplot

```{r}
ggplot(data=penguins,mapping=aes(x=bill_length_mm,y=bill_depth_mm))+
  geom_point()
```

## Line chart

```{r}
ggplot(data=penguins,mapping=aes(x=bill_length_mm,y=bill_depth_mm))+
  geom_line()
```

## Barplot

```{r}
ggplot(data=penguins,
       mapping=aes(x=species,y=body_mass_g))+
  geom_bar()
```

```{r}
ggplot(data=penguins,
       mapping=aes(x=species,y=body_mass_g))+
  geom_bar(stat='identity')
```

## Boxplot

```{r}
ggplot(data=penguins,
       mapping=aes(x=species,y=body_mass_g))+
  geom_boxplot()
```

## Trajectory

```{r}
tracking_data<-GPS_raw
```

```{r}
ggplot(data = tracking_data, aes(x = Latitude, y = Longitude)) +
  geom_path()
```

## Map

```{r}
library(GermanNorthSea)
```

```{r}
Europe<-German_land
```

```{r}
ggplot(data=Europe)+ 
  geom_sf()
```

# ~~~~~~~~~~~~~~~~
# Third Part

# Maps

Identify directory

```{r, message=TRUE, message=FALSE}
My_directory<-here('Downloads')
# My_directory<-"..."
list.files(My_directory)
```

```{r, message=TRUE, message=FALSE}
My_shapefile<-paste0(My_directory,"/Europe.shp")
```

Load shapefile into your environment

```{r}
library(sf)
```

```{r}
Europe_shp<-st_read(My_shapefile)
```

```{r}
ggplot(data = Europe_shp)+  
  geom_sf()
```

## Coords

```{r}
ggplot(data = Europe_shp)+  
  geom_sf()+
  coord_sf(xlim = c(6, 10),ylim = c(53, 56))
```

## Scale

```{r}
library(ggspatial)
```

```{r}
ggplot(data = Europe_shp)+  
  geom_sf()+
  coord_sf(xlim = c(6, 10),ylim = c(53, 56))+
  annotation_scale(location = "br")
```

## Arrow

```{r}
ggplot(data = Europe_shp)+  
  geom_sf()+
  coord_sf(xlim = c(6, 10),ylim = c(53, 56))+
    annotation_scale(location = "br")+
    annotation_north_arrow(location = "tl")
```

```{r}
ggplot(data = Europe_shp)+  
  geom_sf()+
  coord_sf(xlim = c(6, 10),ylim = c(53, 56))+
    annotation_scale(location = "br")+
    annotation_north_arrow(location = "tl", 
                           style = north_arrow_minimal())
```

## CRS

```{r}
Europe_3035<-sf::st_transform(Europe_shp, 3035)
```

```{r}
ggplot(data=Europe_3035)+ 
  geom_sf()+
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000))
```

## Pause

```{r}
ggplot(data = Europe_shp)+  
  geom_sf()+
  coord_sf(xlim = c(6, 10),ylim = c(53, 56))+
    annotation_scale(location = "br")+
    annotation_north_arrow(location = "tl", 
                           style = north_arrow_minimal())
```


# ~~~~~~~~~~~
# Fourth part

# Aesthetics

## Colors

By name

```{r}
ggplot(data = Europe_shp)+  
  geom_sf(color = "orange",
          fill = "yellow")+
  coord_sf(xlim = c(3, 9),ylim = c(53, 56))
```

By hex code

```{r}
ggplot(data = Europe_shp)+  
    geom_sf(color = "#e76f51",
            fill = "#e9c46a")+
      coord_sf(xlim = c(3,9), ylim = c(53,56))
```

## Object

```{r}
My_plot<-ggplot(data = Europe_shp)+  
    geom_sf(color = "#e76f51",
            fill = "#e9c46a")+
      coord_sf(xlim = c(3,9), ylim = c(53,56))
```

```{r}
My_plot
```

## Theme

```{r}
My_plot+
  theme_classic()
```

```{r}
My_plot+
  theme_bw()
```

```{r}
My_plot+
  theme_minimal()
```

```{r}
My_plot+
  theme_void()
```

```{r}
My_plot<-My_plot+
  theme_bw()
My_plot
```

## Background

```{r}
My_plot<-My_plot +
  theme(panel.background = element_rect(fill = '#a8dadc'))
My_plot
```

```{r}
My_plot<-My_plot +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
My_plot
```

## Axis legend

```{r}
My_plot<-My_plot +
  xlab('Longitude')+ylab('Latitude')
My_plot
```

```{r}
My_plot<-My_plot +
    scale_x_continuous(labels = function(x) paste0(x, '\u00B0', "W")) +
    scale_y_continuous(labels = function(x) paste0(x, '\u00B0', "N"))
My_plot
```

## Pause

```{r}
Base_map<-ggplot(data = Europe_shp)+ 
    geom_sf(color = "#e76f51", 
            fill = "#e9c46a")+ 
  coord_sf(xlim = c(3,9), ylim = c(53,56))+
  theme_bw()+
  theme(panel.background = element_rect(fill = '#a8dadc'),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
  xlab('Longitude')+ylab('Latitude')+
  scale_x_continuous(labels = function(x) paste0(x, '\u00B0', "W"))+
  scale_y_continuous(labels = function(x) paste0(x, '\u00B0', "N"))
Base_map
```

# ~~~~~~~~~~~~
# Fifth part
# Add geometries

```{r}
survey_4326<-survey_4326
```

```{r}
head(survey_4326,5)
```

```{r}
Base_map+
  geom_point(data=survey_4326,aes(x = longitude,y= latitude))
```

## Change color 

By name

```{r}
Base_map+
  geom_point(data=survey_4326,aes(x = longitude,y= latitude),
             color='red')
```

By variable

```{r}
Base_map+
  geom_point(data=survey_4326,aes(x = longitude,y= latitude, 
                                  color=month))
```

By taste

```{r}
Base_map+
  geom_point(data=survey_4326,aes(x = longitude,y= latitude, 
                                  color=month))+
  scale_colour_manual(values = c("purple", "yellow", "orange","green","blue"))
```

## Size

```{r}
Base_map+
  geom_point(data=survey_4326,aes(x = longitude,y= latitude),
             size=0.01)
```

## geom_path

```{r}
Base_map+
  geom_path(data=survey_4326,aes(x = longitude,y= latitude))
```

# ~~~~~~~~~~~~
# Sixth part

## Interactive

```{r}
library(plotly)
```

```{r}
ggplotly(Base_map+
  geom_point(data=survey_4326,aes(x = longitude,y= latitude, 
                                  color=date))+
  scale_colour_manual(values = c("purple", "yellow", "orange","green","blue"))
)
```

# Export

## ggsave

Select folder

```{r}
ResultsFolder<-here('Downloads')
```

Give the name of the plot

```{r}
ggsave(My_plot, 
       filename = paste0(ResultsFolder,"/My_plot.png"),
       width = 24, 
       height = 24, 
       units = "in",
       dpi = 500)
```

# End of document

